### 2015-11-07 journal-formats/journal-of-the-royal-society-interface/drones-royal-society.tex

% To compile:
% latex drones
% bibtex drones
% latex drones
% latex drones
% dvipdf drones
\documentclass[12pt]{article}

\usepackage{hyperref}
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

\usepackage[T1]{fontenc}

\usepackage[nofiglist,nofighead,nomarkers]{endfloat}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{sidecap}
\usepackage[pdftex]{graphicx}
\usepackage{caption}
\usepackage{subcaption}
% \usepackage{cite}
\usepackage{color} 
\usepackage{soul}
\usepackage[super,comma,sort&compress]{natbib}
\usepackage{url}
\urlstyle{rm} % typeset urls in roman instead of in typewriter style
\makeatletter
\renewcommand\@biblabel[1]{#1} %from [1] to 1
\makeatother

\usepackage[margin=21mm]{geometry}
\usepackage[labelfont=bf,font=small,labelsep=period,justification=raggedright]{caption}
%% \bibliographystyle{./vancouver}
\bibliographystyle{unsrtabbrv}

\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\pagestyle{myheadings}
\begin{document}

\begin{flushleft}
{\huge
\textbf{\input{drones.title}}
}\smallbreak

Javier Garcia-Bernardo\textsuperscript{1}, 
Peter Sheridan Dodds\textsuperscript{2}, 
Neil F. Johnson\textsuperscript{3}
\\
\textsuperscript{1}Department of Computer Science, University of Vermont, Burlington VT 05405, USA \\
\textsuperscript{2}Computational Story Lab, Department of Mathematics and Statistics, Vermont Complex Systems Center \& the Vermont Advanced Computing Core, University of Vermont, Burlington VT 05405, USA \\
\textsuperscript{3}Department of Physics, University of Miami, Coral Gables, FL 33124, USA
\end{flushleft}

\section*{Abstract}
\input{drones.abs}

\noindent \textbf{Keywords}: Drones, escalation, conflict, patterns

\input{drones.body}

\bibliography{drones}

\section*{Acknowledgements}
\input{drones.acknowledgments}

\bigskip
\bigskip

\noindent
\textbf{Short title:} \input{drones.title}

\end{document}