### 2023-03-16 correspondence/2022-05-31response-epj-data-science.outtakes.tex

\todo{Zipf distributions comparing cities in 1750 to 2000 in the US}

\todo{Could compare normalized ones ...}

\todo{Bad idea: Joint distribution means that they won't be sensible
  Zipf distributions with the 0 point. Plus binning.
  Best thing would be to overlay the two Zipf distributions. We
  considered ways of adding Zipf distributions to the plots but we
  felt that the results would be too complicated (and we acknowledge
  that the allotaxonographs are complicated enough as is)}

\noway{we have also expanded the balance section to included absolute numbers}.