### 2025-06-26 paper-template-newpax.tex
General updates

\documentclass{article}

%% load the lua code
\directlua{require("newpax")}

%% write .newpax files for newpax.sty

%% for references to the main data set
\newcommand{\Ncharacters}{2000}
\newcommand{\Ntraits}{464}
\newcommand{\Nstories}{341}

%%   newpax.writenewpax("figures/localized/Arrested-Development-Lucille-Bluth-\Ncharacters-\Ntraits-\Nstories")
%%   newpax.writenewpax("figures/localized/Buffy-the-Vampire-Slayer-Buffy-Summers-\Ncharacters-\Ntraits-\Nstories")
%%   newpax.writenewpax("figures/localized/rational--whimsical-\Ncharacters-\Ntraits-\Nstories")

\directlua
    {
      newpax.writenewpax("figures/localized/stories-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/traits-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/archetypes-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/essential-traits-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/orderly--chaotic-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Its-Always-Sunny-in-Philadelphia-Charlie-Kelly-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Buffy-the-Vampire-Slayer-Buffy-Summers-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Harry-Potter-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Harry-Potter-dimensions-3-vs-4-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/comparison-Archer-Malory-Archer-vs-Arrested-Development-Lucille-Bluth-2000-464-341")
      newpax.writenewpax("figures/localized/comparison-Archer-Sterling-Archer-vs-Bobs-Burgers-Bob-Belcher-2000-464-341")
    }

\begin{document}
\end{document}

%% - these don't work properly because the cropped pdfs include all of the original
%% - moving to pngs
%% 
%%       newpax.writenewpax("figures/localized/orderly--chaotic-\Ncharacters-\Ntraits-\Nstories")
%%       newpax.writenewpax("figures/localized/orderly--chaotic-\Ncharacters-\Ntraits-\Nstories-title")
%%       newpax.writenewpax("figures/localized/Buffy-the-Vampire-Slayer-Buffy-Summers-\Ncharacters-\Ntraits-\Nstories-title")
%%       newpax.writenewpax("figures/localized/outcast-angel-hero-header")
%%       newpax.writenewpax("figures/localized/outcast-angel-hero-switcher")
%%       newpax.writenewpax("figures/localized/essential-trait-007-positive-header")